Storm Surges

### 2014-07-15 RoughWork/manuscript.tex
Modified feb2006 by removing a background state.

\section{Storm Surge Hindcasts}\label{sec:storm}
%outline: 
%1. factors contributing to storm surge with focus on Feb 2006 event
%2. Strong wind event but no surge (maybe Dec 2006 or Nov 2009)
%3. High resolution wind (hopefully Dec 2012)
A series of storm surge hindcasts is presented next. Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents. Stratification conditions from model spin-up at a time close the the event (within 5 days) were chosen to initialize the temperature and salinity fields. Boundary conditions include atmospheric forcing and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents. Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height predictions due to the omission of the next leading order tidal constituents such as J$_1$. Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm (check) during times of high tidal range, and in particular, during the storm surge season Nov-Feb. As such, storm surge model predictions were corrected by accounting for this error. Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height.

@@

Three significiant surge or wind events have been considered. First, a study of a large surge event on Feb 4, 2006 is presented. This event caused siginificant damage and flooding in Boundary Bay, a community south of Richmond. (citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hindcasting. 
%% Trying to decide Dec 15, 2005 or Nov 2009. I think it's more sense to do Nov 2009 - strong winds but no surge. 
Next, an extreme wind event which resulted in property damage in Vancouver and surrounding areas is discussed. This event occured on Dec 15, 2006 but did not result in significant flooding such as that which occured on Feb 4, 2006 because the storm did not occur at high tide. This case was choosen to identify the model's capability in handling large wind events that do not result in high water levels. Finally, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.

@@

The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations and corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River. The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides. There is also a good match in amplitude of the water level maximum at Point Atkinson and Campbell River, with a relative error less than 2\%. Indeed, there is only a 1.2 cm difference between observed and modelled water level elevation at Point Atkinson and a 9 cm difference at Campbell River. However, the error at Victoria and Patricia Bay is more siginificant, reaching 12\% and 8\% respectively (close to 40 cm in each case), partly due to error in tidal amplitudes as discussed above. The precise location and magnitude of the M$_2$ amphidrome around Victoria is very sensitive to model parameters such as bottom friction.

The error in modelled residuals at Victoria and Patricia Bay is also apparent in the panels on the right, most noticeable is an underestimate in modelled residual throughout most of the time period. Additionally, there is a delay of about 2-3 hours in the timing of maximum residual at all four locations. The errors in maximum residual range from 3.9 cm at Point Atkinson to 21 cm at Patricia Bay. 
%Question: is it better to present relative errors or just difference?

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/feb2006.pdf}
\caption{Comparison of observations and model output for the Feb 4, 2006 storm surge. Left: total water level observation (solid) and corrected model (dashed). Right: observed residuals (solid) and modelled residuals (dashed).}
\label{fig:feb2006}
\end{figure}

>>>>

\section{Storm Surge Hindcasts}\label{sec:storm}
%outline: 
%1. factors contributing to storm surge with focus on Feb 2006 event
%2. Strong wind event but no surge (maybe Dec 2006 or Nov 2009)
%3. High resolution wind (hopefully Dec 2012)
A series of storm surge hindcasts is presented next. Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents. Stratification conditions from model spin-up at a time close to the event (within 5 days) were chosen to initialize the temperature and salinity fields. Boundary conditions include atmospheric forcing and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents. Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height predictions due to the omission of the next leading order tidal constituents such as J$_1$. Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm (check) during times of high tidal range, and in particular, during the storm surge season Nov-Feb. As such, storm surge model predictions were corrected by accounting for this error. Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height and removing a background anomaly caused by topographic pressure gradients since the atmospheric reforecasting model uses a terrain following coordinate system. (Nancy: I think we need a better way of adjusting come publication time).

@@

Three significiant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure. First, a study of a large surge event on Feb 4, 2006 is presented. This event caused siginificant damage and flooding in Boundary Bay, a community south of Richmond. (citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hindcasting. 
%% Trying to decide Dec 15, 2005 or Nov 2009. I think it's more sense to do Nov 2009 - strong winds but no surge. 
Second, an extreme wind event which resulted in property damage in Vancouver and surrounding areas is discussed. This event occured on Dec 15, 2006 but did not result in significant flooding such as that which occured on Feb 4, 2006 because the storm did not occur at high tide. This case was choosen to identify the model's capability in handling large wind events that do not result in high water levels. Third, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed. Lastly, these cases presented in detail will be combined with $x$ additional runs for statistical evaluation of model performance.

@@

The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations and corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River. The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides. There is also a good match in amplitude of the water level maximum at Point Atkinson and Campbell River, with a relative error less than 4\%. There is only a 7.2 cm difference between observed and modelled water level elevation at Point Atkinson and a 20.9 cm difference at Campbell River. Tthe error at Victoria and Patricia Bay is slightly more siginificant, reaching 9.6\% and 6.3\% respectively (close to 30 cm in each case), partly due to error in tidal amplitudes as discussed above. The precise location and magnitude of the M$_2$ amphidrome around Victoria is very sensitive to model parameters such as bottom friction.

The error in modelled residuals is also apparent in the panels on the right, most noticeable is an underestimate in maximum residual (or surge) at all locations. There is also a 2-3 hour delay in the timing of the maximum residual except at Victoria where the maximum residual in sync with observations. The errors in maximum residual range from 8.0 cm at Victoria to 28 cm at Campbell River. It is also apparent that Point Atkinson and Campbell River experienced the largest observed surges at 82 cm and 95 cm respectively. The surges at Victoria and Patricia Bay were observed to be 66 cm and 74 cm. 
%Question: is it better to present relative errors or just difference?

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/feb2006_bg.pdf}
\caption{Comparison of observations and model output for the Feb 4, 2006 storm surge. Left: total water level observation (solid) and corrected model (dashed). Right: observed residuals (solid) and modelled residuals (dashed).}
\label{fig:feb2006}
\end{figure}
